Non-contextual Hidden Variable Model for Quantum Mechanics

### 2017-12-17 version2postPRL/nonContextual.tex
Major update after the PRL feedback and reading the book, papers etc.

\documentclass[british,aps,prl,superscriptaddress,nofootinbib,times,reprint]{revtex4-1}
%\documentclass[british,aps,superscriptaddress,nofootinbib,twocolumn,prl]{revtex4-1}
%\usepackage[T1]{fontenc}
%\usepackage[latin9]{inputenc}
\usepackage{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm,headheight=2cm,headsep=2cm,footskip=2cm}
\setcounter{secnumdepth}{3}
\usepackage{verbatim}
\usepackage{refstyle}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
%\usepackage{esint}
\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\AtBeginDocument{\providecommand\thmref[1]{\ref{thm:#1}}}
\AtBeginDocument{\providecommand\lemref[1]{\ref{lem:#1}}}
\AtBeginDocument{\providecommand\propref[1]{\ref{prop:#1}}}
\AtBeginDocument{\providecommand\eqref[1]{\ref{eq:#1}}}
\AtBeginDocument{\providecommand\figref[1]{\ref{fig:#1}}}
\AtBeginDocument{\providecommand\secref[1]{\ref{sec:#1}}}
\AtBeginDocument{\providecommand\ssecref[1]{\ref{ssec:#1}}}
\AtBeginDocument{\providecommand\defnref[1]{\ref{defn:#1}}}
\AtBeginDocument{\providecommand\tblref[1]{\ref{tbl:#1}}}

@@

@@

\title{A non-contextual hidden variable model for quantum mechanics}
\author{Atul Singh Arora}
\email{ms11003@iisermohali.ac.in}
%\selectlanguage{british}%
\affiliation{Department of Physical Sciences, 
Indian Institute of Science Education and Research
Mohali, Sector 81 SAS Nagar 140306 Punjab India.}
\author{Arvind}
\email{arvind@iisermohali.ac.in}
\affiliation{Department of Physical Sciences, 
Indian Institute of Science Education and Research
Mohali, Sector 81 SAS Nagar 140306 Punjab India.}
\author{Kishor Bharti}
\affiliation{Centre for Quantum Technologies,
National University of Singapore}
%\selectlanguage{british}%
\begin{abstract}
We propose a non-contextual hidden variable model,
consistent with all kinematic
predictions of quantum
mechanics (QM).  A careful scrutiny of consistency
requirements of any hidden variable model with
quantum mechanics and the corresponding no-go
theorems leads us to the conclusion that,
contrary to common belief,  the notion of
contextuality is not a necessary feature of quantum kinematics (QK).
The alternative view that emerges, hinges on
identifying a new classical property which we call
``multiplicativity''.  It turns out that by
relaxing this condition of ``multiplicativity'' on
hidden variable models, they can be made
consistent with QK while remaining
non-contextual.  The connection of
non-``multiplicativity'' to Bohm's model and to
apparent contextuality is also discussed.
\end{abstract}
\pacs{}

@@

@@

However, not all HV models (e.g.

@@

trajectories~\cite{Bohm1,Bohm2}) are incompatible

@@

\begin{figure}[h]
\includegraphics[width=0.9\columnwidth]{block1}
\caption{Non-contextuality is not inconsistent
with kinematic predictions of QM} \label{fig:block}\end{figure}

@@

QM (see \figref{block}).

@@

to play in QIP protocols.  This provides a new

@@

\begin{notn*}
(a) $\psi\in\mathcal{H}$ represents a pure quantum
mechanical state of the system in the Hilbert
space $\mathcal{H}$, (b) $\hat{\mathcal{H}}$ is
defined to mean the set of Hermitian observables
for the system, (c) $[\mathcal{H}]$ is defined to
mean $\{\mathcal{H},\,\mathbb{R}^{\otimes}\}$, which
represents the state  space of the
system including HVs, (d) $[\psi]\in[\mathcal{H}]$
will represent the state of the system including
HVs, (e) a prediction map is $M:\{
\hat{\mathcal{H}},[\mathcal{H}] \}\to\mathbb{R}$,
(f) a sequence map is $S:\{
\hat{\mathcal{H}},[\mathcal{H}],\mathbb{R}
\}\to[\mathcal{H}]$, (g) $f$ is an arbitrary map
from $\{
\hat{\mathcal{H}},\hat{\mathcal{H}},\dots\hat{\mathcal{H}}
\} \to \hat{\mathcal{H}}$ constructed using
multiplication and addition of compatible
observables, and multiplication with complex
numbers, (h) $\tilde{f}$ is a map constructed by
replacing observables in $f$ with real numbers.
\end{notn*}
\begin{defn} A theory is non-contextual, if it
provides a map $M: \{
\hat{\mathcal{H}},[\mathcal{H}] \} \to\mathbb{R}$
to explain measurement outcomes. A theory which is
not non-contextual is contextual~\cite{peresBook}.
\end{defn}

@@

\begin{defnrem}
Broader definitions in the literature have been
suggested which declare a larger set of theories
as non-contextual~\cite{spekkens_pra}. For our
purposes, however, this restricted definition will
suffice  and our results will in any case be valid
for the broader definitions. The term contextual
is used to suggest that the value an observable
takes might depend on which other compatible
observable it is being  measured with.
\end{defnrem}
Before we re-examine the result of Kochen and
Specker~\cite{KochenSpecker} that non-contextual
models cannot be compatible with quantum
mechanics and show that multiplicativity can provide an
alternative to contextuality, we define
multiplicativity explicitly.
\begin{defn}
A prediction map $M$ is \emph{multiplicative} iff
\begin{align*} &
M(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi])
= \\ & \tilde
f(M(\hat{B}_{1},[\psi]),M(\hat{B}_{2},[\psi]),\dots
M(\hat{B}_{N},[\psi])), \end{align*}
where $\hat{B}_{i}\in\hat{\mathcal{H}}$ are
arbitrary mutually commuting observables and
$[\psi]\in[\mathcal{H}]$. A
\emph{non-multiplicative} map is one that is not
multiplicative.  
\end{defn} 
Note that if $M$ is taken to represent the
measurement outcome (in QM), then for states of
the system which are simultaneous eigenkets of
$\hat{B}_{i}$s, $M$ must clearly be
multiplicative. It is, however, not obvious that
this property must always hold. For example,
consider two spin-half particles in the state 
$\left|1\right\rangle \otimes\left|1\right\rangle
$ written in the computational basis and 
operators 
$\hat{B}_{1}=\hat{\sigma}_{x}\otimes\hat{\sigma}_{x}$,
$\hat{B}_{2}=\hat{\sigma}_{y}\otimes\hat{\sigma}_{y}$
and
$\hat{C}=\hat{B}_{1}\hat{B}_{2}=-\hat{\sigma}_{z}\otimes\hat{\sigma}_{z}$
written in terms of Pauli operators.
We must have $M(\hat{C})=-1$ while
$M(\hat{B}_{1})=\pm1$ and $M(\hat{B}_{2})=\pm1$
independently, according to QM, with probability
half. Here multiplicativity clearly is not
required to hold.
Antithetically, it is clear that if one first
measures $\hat{B}_{1}$ and subsequently measures
$\hat{B}_{2}$, then the product of the results
must be $-1$. This is consistent with measuring
$\hat{C}$. To capture this property we define
\emph{sequential multiplicativity} as follows.
\begin{defn} 
A prediction map $m$ is
\emph{sequentially multiplicative} for a given
sequence map $s$, iff \begin{align*}
&M(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi_{1}])=\\
&\tilde
f(M(\hat{B}_{1},[\psi_{k_{1}}]),M(\hat{B}_{2},[\psi_{k_{2}}]),\dots,M(\hat{B}_{N},[\psi_{k_{N}}])),
\end{align*} 
where
$\{k_{1},k_{2},\dots,k_{N}\} \in\{
N! {\rm ~permutations~of~}  k'{\rm s} \}$, 
$\hat{B}_{i}\in\hat{\mathcal{H}}$ are
arbitrary mutually commuting observables,
$[\psi_{i}]\in[\mathcal{H}]$ and
$[\psi_{k+1}] :=
S(\hat{B}_{k},[\psi_{k}],M(\hat{B}_{k},[\psi_{k}]))$,
$\forall\,[\psi_{i}]$.  
\label{defn:seqnMltpl}
\end{defn} 
With these definitions we are now ready to discuss
the `proof of contextuality'. We first state the
contextuality theorem in our notation:
\begin{thm*} Let a map
$M:\hat{\mathcal{H}}\to\mathbb{R}$, be s.t. (a)
$M(\hat{\mathbb{I}})=1$, (b)
$M(f(\hat{B}_{1},\hat{B}_{2},\dots))=\tilde
f(M(\hat{B}_{1}),M(\hat{B}_{2}),\dots)$, for any
arbitrary function $f$, where $\hat{B}_{i}$ are
mutually commuting Hermitian operators. If $m$ is
assumed to describe the outcomes of measurements,
then no $M$ exists which is consistent with all
predictions of QM. 
\label{thm:KS}
\end{thm*}

\begin{proof} Peres Mermin [PM]
($\left|\mathcal{H}\right|\ge4$)
\cite{Peres,Mermin}:
For a system composed of two spin-half particles 
consider the following set of operators \[
\hat{A}_{ij}\doteq\left[\begin{array}{ccc}
\hat{\mathbb{I}}\otimes\hat{\sigma}_{x} &
\hat{\sigma}_{x}\otimes\hat{\mathbb{I}} &
\hat{\sigma}_{x}\otimes\hat{\sigma}_{x}\\
\hat{\sigma}_{y}\otimes\hat{\mathbb{I}} &
\hat{\mathbb{I}}\otimes\hat{\sigma}_{y} &
\hat{\sigma}_{y}\otimes\hat{\sigma}_{y}\\
\hat{\sigma}_{y}\otimes\hat{\sigma}_{x} &
\hat{\sigma}_{x}\otimes\hat{\sigma}_{y} &
\hat{\sigma}_{z}\otimes\hat{\sigma}_{z}
\end{array}\right] \] which have the property that
all operators along a row (or column) commute. Further,
the product of rows (or columns) yields
$\hat{R}_{i}=\hat{\mathbb{I}}$ and
$\hat{C}_{j}=\hat{\mathbb{I}}\,(j\neq3)$,
$\hat{C}_{3}=-\hat{\mathbb{I}}$, ($\forall\,i,j$) where
$\hat{R}_{i}:=\prod_{j}\hat{A}_{ij}$,
$\hat{C}_{j}:=\prod_{i}\hat{A}_{ij}$. Let us
assume that $M$ exists. From property (b) of the
map, to get $M(\hat{C}_{3})=-1$ (as required by
property (a)), we must have an odd number of $-1$
assignments in the third column. In the remaining
columns, the number of $-1$ assignments must be
even (for each column). Thus, in the entire
square, the number of $-1$ assignments must be
odd. Let us use the same reasoning, but along the
rows. Since each $M(\hat{R}_{i})=1$, we must have
even number of $-1$ assignments along each row.
Thus, in the entire square, the number of $-1$
assignments must be even. We have arrived at a
contradiction and therefore we conclude 
that $M$ does not exist
\end{proof}

\begin{rem} One could in principle assume $M$, to
be s.t. (a) $M(\hat{\mathbb{I}})=1$, (b)
$M(\alpha\hat{B}_{i})=\alpha M(\hat{B}_{i})$, for
$\alpha\in\mathbb{R}$, (c)
$M(\hat{B}_{i}^{2})=M(\hat{B}_{i})^{2}$, (d)
$M(\hat{B}_{i}+\hat{B}_{j})=M(\hat{B}_{i})+M(\hat{B}_{j})$,
to deduce (d)
$M(\hat{B}_{i}\hat{B}_{j})=M(\hat{B}_{i})M(\hat{B}_{j})$
and that $M(\hat{B}_{i})\in$ spectrum of
$\hat{B}_{i}$. Effectively then, condition (b)
listed in the theorem is satisfied as a
consequence.  Therefore, assuming (a)-(d) as
listed above, rules out a larger class of
$M$~\cite{KochenSpecker}.
\end{rem} 
Here $M$ maybe viewed as a specific class of
prediction maps, that implicitly depends on the
state $[\psi]$. It is clear that according to the
theorem, non-contextual maps which are
\emph{multiplicative} must be incompatible with
QM. 
This leaves open an interesting possibility
that non-contextual maps which are
non-\emph{multiplicative} could be consistent with
QM. We explicitly construct such an HV model and use
\emph{sequential multiplicativity} as a consistency
check imposed by QM.
\begin{prop*} 
Let a quantum mechanical 
system be in a state, s.t. measurement of
$\hat{C}$ yields repeatable results (same result
each time). Then according to QM, \emph{sequential
multiplicativity} holds, where $\hat{C}:=
f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n})$, and
$\hat{B}_{i}$ are as defined (in \defnref{seqnMltpl})
\end{prop*}
\begin{proof} 
Without loss of generality we can take
$\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}$ to be
mutually compatible and a complete set of
observables (operators can be added to make the set
complete if needed).
It follows
that $\exists$
$\left|\mathbf{b}=\left(b_{1},b_{2},\dots
b_{n}\right)\right\rangle $ s.t.
$\hat{B}_{i}\left|\mathbf{b}\right\rangle
=b_{i}\left|\mathbf{b}\right\rangle $,
and that
$\sum_{\mathbf{b}}\left|\mathbf{b}\right\rangle
\left\langle \mathbf{b}\right|=\hat{\mathbb{I}}$.
Let the state of the system
$\left|\psi\right\rangle $ be s.t.
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $. For
the statement to follow, one need only show that
$\left|\psi\right\rangle $ must be made of only
those $\left|\mathbf{b}\right\rangle $s, which
satisfy $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$.  This is the crucial step and
proving this is straightforward. We start with
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $ and take its inner
product with $\left\langle \mathbf{b}\right|$ to
get 
\begin{eqnarray*} 
c \langle \mathbf{b}|\psi\rangle 
& = &
\langle
\mathbf{b} |\hat{C} |\psi
\rangle  
\\
&=&
\langle
\mathbf{b}| f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}) |
\psi \rangle
\\ 
&=&
\tilde f (b_{1},b_{2},\dots b_{n})
\langle
\mathbf{b}|\psi\rangle  
\end{eqnarray*}
Also, we have $\left|\psi\right\rangle
=\sum_{\mathbf{b}}\left\langle
\mathbf{b}|\psi\right\rangle
\left|\mathbf{b}\right\rangle $, from
completeness. If we consider
$\left|\mathbf{b}\right\rangle $s for which
$\left\langle \mathbf{b}|\psi\right\rangle \neq0$,
then we find that indeed $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$. The case when $\left\langle
\mathbf{b}|\psi\right\rangle =0$ is anyway
irrelevant as the corresponding $\vert b\rangle$
does not contribute to $\vert \psi\rangle$.
We can thus conclude that
$\left|\psi\right\rangle $ is made only of those
$\left|\mathbf{b}\right\rangle $s that satisfy the
required relation.
\end{proof} 
It is worth noting that in the Peres Mermin
case, where $\hat{R}_{i}$ and $\hat{C}_{j}$ are
$\pm\hat{\mathbb{I}}$, it follows that all
states are their eigenstates. Consequently, for
these operators \emph{sequential multiplicativity}
must always hold.
%\subsection{C-ingle model | An Explicit
%Construction \label{ssec:cingle}}

>>>>

\documentclass[doublecol,british]{epl2} 
% or \documentclass[page-classic]{epl2} for one column style
\usepackage[normalem]{ulem}

%\setcounter{secnumdepth}{3}     
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{refstyle}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{geometry}
\geometry{verbose,tmargin=3cm,bmargin=2cm,lmargin=2cm,rmargin=2cm,headheight=2cm,headsep=2cm,footskip=1cm}
\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\AtBeginDocument{\providecommand\thmref[1]{\ref{thm:#1}}}
\AtBeginDocument{\providecommand\lemref[1]{\ref{lem:#1}}}
\AtBeginDocument{\providecommand\propref[1]{\ref{prop:#1}}}
\AtBeginDocument{\providecommand\eqref[1]{\ref{eq:#1}}}
\AtBeginDocument{\providecommand\figref[1]{\ref{fig:#1}}}
\AtBeginDocument{\providecommand\secref[1]{\ref{sec:#1}}}
\AtBeginDocument{\providecommand\ssecref[1]{\ref{ssec:#1}}}
\AtBeginDocument{\providecommand\defnref[1]{\ref{defn:#1}}}
\AtBeginDocument{\providecommand\tblref[1]{\ref{tbl:#1}}}

@@

%%%%%%%%%%%%%%%%%%%

\title{Non-contextual Hidden Variable Model for Quantum Mechanics}
\shorttitle{Non-contextual HV Model for QM} %Insert here a short version of the title if it exceeds 70 characters

\author{Atul Singh Arora\inst{1,2} \and Arvind\inst{1} \and Kishor Bharti\inst{3}}
\shortauthor{A. S. Arora \etal}

\institute{                    
  \inst{1} Indian Institute of Science Education and Research (IISER), Mohali,\\Sector 81 SAS Nagar 140306 Punjab India.\\
  \inst{2} Université libre de Bruxelles, \\50 av. F. D. Roosevelt - CP 165/59, B-1050 Brussels, Belgium.\\
  \inst{3} Centre for Quantum Technologies (CQT), \\National University of Singapore (NUS) -  Block S15, 3 Science Drive 2, Singapore 117543.
}

\pacs{03.65.Ta}{Foundations of quantum mechanics}
%\pacs{}{}
% \pacs{nn.mm.xx}{First pacs description}
% \pacs{nn.mm.xx}{Second pacs description}
% \pacs{nn.mm.xx}{Third pacs description}

\abstract{ \color{blue}
We construct a simple non-contextual hidden variable model consistent with all kinematic predictions of quantum mechanics (QM). The famous Bell-KS theorem shows that non-contextual models which satisfy a further reasonable restriction are inconsistent with QM. We define a weaker variant of this restriction which captures its essence while still allowing a non-contextual description of QM. This is in contrast to contextual toy hidden variable models, such as the one by Bell, and brings out an interesting alternate way of looking at QM. This is also related to Bohm's model where it is harder to pin down such features.}

@@

@@

\section{Introduction}

@@

{\color{blue} While there have been generalisations, in this letter, we restrict ourselves to the standard notion of non-contextuality as used by KS [??].} 
Not all HV models (e.g.

@@

trajectories~\cite{Bohm1,Bohm2}), however, are incompatible

@@

% \begin{figure}[h]
% \includegraphics[width=0.9\columnwidth]{block1}
% \caption{Non-contextuality is not inconsistent
% with kinematic predictions of QM} \label{fig:block}\end{figure}
{\color{blue}
The KS theorem is applicable to non-contextual HV models which additionally satisfy functional consistency --- algebraic constraints obeyed by commuting quantum observables must be satisfied by the HV model at the level of individual outcomes. One of the main justifications given for imposing this was the reasonable requirement that an observable $\hat A^2$ must depend on $\hat A$ even at the HV level because otherwise one can construct HV models where these observables get mapped to random variables with no relation to each other (see [??]). As we will show, one can weaken this requirement into what we call weak functional consistency --- functional consistency is demanded only when the observables in question have sharp values. This entails that even in the HV model the observables will depend on each other where they should thereby capturing the same idea. The consequence of demanding only the weaker variant is that the KS theorem no longer applies and non-contextual HV models can be consistently constructed. This is in stark contrast with contextual HV models, such as the one by Bell, which we demonstrate by re-examining a `proof of contextuality' using our model. 
}
\sout{

@@

QM  %(see \figref{block}).

@@

to play in QIP protocols. }
This provides a new

@@

\section{Non-Contextuality and Functional Consistency}
We introduce some notation and make the relevant notions precise to facilitate the description and assesment of our model.
\begin{notn*}
(a) $\psi\in\mathcal{H}$ represents a pure quantum
mechanical state of the system in the Hilbert
space $\mathcal{H}$, (b) $\hat{\mathcal{H}}$ is
defined to mean the set of Hermitian observables
for the system, (c) $[\mathcal{H}]$ is defined to
mean $\{\mathcal{H},\,\mathbb{R}^{\otimes}\}$, which
represents the state  space of the
system including HVs, (d) $[\psi]\in[\mathcal{H}]$
will represent the state of the system including
HVs, (e) a prediction map is $M:\{
\hat{\mathcal{H}},[\mathcal{H}] \}\to\mathbb{R}$,
(f) a sequence map is $S:\{
\hat{\mathcal{H}},[\mathcal{H}],\mathbb{R}
\}\to[\mathcal{H}]$, (g) $f$ is an arbitrary map
from $\{
\hat{\mathcal{H}},\hat{\mathcal{H}},\dots\hat{\mathcal{H}}
\} \to \hat{\mathcal{H}}$ constructed using
multiplication and addition of compatible
observables, and multiplication with complex
numbers, (h) $\tilde{f}$ is a map constructed by
replacing observables in $f$ with real numbers.
\end{notn*}
\begin{defn} A theory is non-contextual, if it
provides a map $M: \{
\hat{\mathcal{H}},[\mathcal{H}] \} \to\mathbb{R}$
to explain measurement outcomes. A theory which is
not non-contextual is contextual~\cite{peresBook}.
\end{defn}

@@

\begin{defnrem}
{\color{blue} Broader definitions in the literature have been suggested~\cite{spekkens_pra} which extend the notion to probabilistic HV models. The idea is that any feature of a HV model that is not determined solely by the operational aspect of QM is defined to be, in the case of measurements for instance, measurement contextuality. To maintain a distinction between different features of HV models, which is of interest here, we stick to the standard definition.}
\sout{
Broader definitions in the literature have been
suggested which declare a larger set of theories
as non-contextual. For our %~\cite{spekkens_pra}. For our
purposes, however, this restricted definition will
suffice  and our results will in any case be valid
for the broader definitions. The term contextual
is used to suggest that the value an observable
takes might depend on which other compatible
observable it is being  measured with. }
\end{defnrem}

{\color{blue} 
In addition to a HV model being non-contextual Kochen and Specker~\cite{KochenSpecker} demand the following to establish their no-go theorem.
}

\sout{
Before we re-examine the result of Kochen and
Specker that non-contextual%~\cite{KochenSpecker} that non-contextual
models cannot be compatible with quantum
mechanics and show that multiplicativity can provide an
alternative to contextuality, we define
multiplicativity explicitly.}

\begin{defn}
A prediction map $M$ is \emph{functionally consistent} iff
\begin{align*} &
M(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi])
= \\ & \tilde
f(M(\hat{B}_{1},[\psi]),M(\hat{B}_{2},[\psi]),\dots
M(\hat{B}_{N},[\psi])), \end{align*}
where $\hat{B}_{i}\in\hat{\mathcal{H}}$ are
arbitrary mutually commuting observables and
$[\psi]\in[\mathcal{H}]$. A
\emph{functionally inconsistent} map is one that is not
functionally consistent.  
\end{defn} 
Note that if $M$ is taken to represent the
measurement outcome (in QM), then for states of
the system which are simultaneous eigenkets of
$\hat{B}_{i}$s, $M$ must clearly be
functionally consistent. It is, however, not obvious that
this property must always hold. For example,
consider two spin-half particles in the state 
$\left|1\right\rangle \otimes\left|1\right\rangle
$ written in the computational basis and 
operators 
$\hat{B}_{1}=\hat{\sigma}_{x}\otimes\hat{\sigma}_{x}$,
$\hat{B}_{2}=\hat{\sigma}_{y}\otimes\hat{\sigma}_{y}$
and
$\hat{C}=\hat{B}_{1}\hat{B}_{2}=-\hat{\sigma}_{z}\otimes\hat{\sigma}_{z}$
written in terms of Pauli operators.
We must have $M(\hat{C})=-1$ while
$M(\hat{B}_{1})=\pm1$ and $M(\hat{B}_{2})=\pm1$
independently, according to QM, with probability
half. Here functional consistency clearly is not
required to hold {\color{blue} and indeed this is why it was demanded in addition to being consistent with QM by KS}.
Antithetically, it is clear that if one first
measures $\hat{B}_{1}$ and subsequently measures
$\hat{B}_{2}$, then the product of the results
must be $-1$. This is consistent with measuring
$\hat{C}$. {\color{blue} In our treatment, instead of imposing functional consistency, we demand its aforesaid weaker variant. It captures the same idea, however, only when it has a precise meaning according to QM.} To capture this property precisely we define
\emph{weak functional consistency} as follows.
\begin{defn} 
A prediction map $m$ has
\emph{weak functional consistency} for a given
sequence map $s$, iff \begin{align*}
&M(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi_{1}])=\\
&\tilde
f(M(\hat{B}_{1},[\psi_{k_{1}}]),M(\hat{B}_{2},[\psi_{k_{2}}]),\dots,M(\hat{B}_{N},[\psi_{k_{N}}])),
\end{align*} 
where
$\{k_{1},k_{2},\dots,k_{N}\} \in\{
N! {\rm ~permutations~of~}  k'{\rm s} \}$, 
$\hat{B}_{i}\in\hat{\mathcal{H}}$ are
arbitrary mutually commuting observables,
$[\psi_{i}]\in[\mathcal{H}]$ and
$[\psi_{k+1}] :=
S(\hat{B}_{k},[\psi_{k}],M(\hat{B}_{k},[\psi_{k}]))$,
$\forall\,[\psi_{i}]$.  
\label{defn:seqnMltpl}
\end{defn}

With these definitions we are now ready to discuss
the `proof of contextuality'. We first state the
contextuality theorem in our notation:
\begin{thm*} Let a map
$M:\hat{\mathcal{H}}\to\mathbb{R}$, be s.t. (a)
$M(\hat{\mathbb{I}})=1$, (b)
$M(f(\hat{B}_{1},\hat{B}_{2},\dots))=\tilde
f(M(\hat{B}_{1}),M(\hat{B}_{2}),\dots)$, for any
arbitrary function $f$, where $\hat{B}_{i}$ are
mutually commuting Hermitian operators. If $m$ is
assumed to describe the outcomes of measurements,
then no $M$ exists which is consistent with all
predictions of QM. 
\label{thm:KS}
\end{thm*}

\begin{proof} Peres Mermin [PM]
($\left|\mathcal{H}\right|\ge4$)
\cite{Peres,Mermin}:
For a system composed of two spin-half particles 
consider the following set of operators 
\begin{equation*}
\hat{A}_{ij}\doteq\left[\begin{array}{ccc}
\hat{\mathbb{I}}\otimes\hat{\sigma}_{x} &
\hat{\sigma}_{x}\otimes\hat{\mathbb{I}} &
\hat{\sigma}_{x}\otimes\hat{\sigma}_{x}\\
\hat{\sigma}_{y}\otimes\hat{\mathbb{I}} &
\hat{\mathbb{I}}\otimes\hat{\sigma}_{y} &
\hat{\sigma}_{y}\otimes\hat{\sigma}_{y}\\
\hat{\sigma}_{y}\otimes\hat{\sigma}_{x} &
\hat{\sigma}_{x}\otimes\hat{\sigma}_{y} &
\hat{\sigma}_{z}\otimes\hat{\sigma}_{z}
\end{array}\right]
\end{equation*}
which have the property that
all operators along a row (or column) commute. Further,
the product of rows (or columns) yields
$\hat{R}_{i}=\hat{\mathbb{I}}$ and
$\hat{C}_{j}=\hat{\mathbb{I}}\,(j\neq3)$,
$\hat{C}_{3}=-\hat{\mathbb{I}}$, ($\forall\,i,j$) where
$\hat{R}_{i}:=\prod_{j}\hat{A}_{ij}$,
$\hat{C}_{j}:=\prod_{i}\hat{A}_{ij}$. Let us
assume that $M$ exists. From property (b) of the
map, to get $M(\hat{C}_{3})=-1$ (as required by
property (a)), we must have an odd number of $-1$
assignments in the third column. In the remaining
columns, the number of $-1$ assignments must be
even (for each column). Thus, in the entire
square, the number of $-1$ assignments must be
odd. Let us use the same reasoning, but along the
rows. Since each $M(\hat{R}_{i})=1$, we must have
even number of $-1$ assignments along each row.
Thus, in the entire square, the number of $-1$
assignments must be even. We have arrived at a
contradiction and therefore we conclude 
that $M$ does not exist
\end{proof}

\begin{rem} One could in principle assume $M$, to
be s.t. (a) $M(\hat{\mathbb{I}})=1$, (b)
$M(\alpha\hat{B}_{i})=\alpha M(\hat{B}_{i})$, for
$\alpha\in\mathbb{R}$, (c)
$M(\hat{B}_{i}^{2})=M(\hat{B}_{i})^{2}$, (d)
$M(\hat{B}_{i}+\hat{B}_{j})=M(\hat{B}_{i})+M(\hat{B}_{j})$,
to deduce (d)
$M(\hat{B}_{i}\hat{B}_{j})=M(\hat{B}_{i})M(\hat{B}_{j})$
and that $M(\hat{B}_{i})\in$ spectrum of
$\hat{B}_{i}$. Effectively then, condition (b)
listed in the theorem is satisfied as a
consequence.  Therefore, assuming (a)-(d) as
listed above, rules out a larger class of
$M$~\cite{KochenSpecker}.
\end{rem} 
Here $M$ maybe viewed as a specific class of
prediction maps, that implicitly depends on the
state $[\psi]$. It is clear that according to the
theorem, non-contextual maps which are
\emph{functionally consistent} must be incompatible with
QM. 
This leaves open an interesting possibility that non-contextual maps which have \emph{weak functional consistency} could be consistent with QM.
{ \color{blue} Before proceeding to do so explicitly we observe that \emph{weak functional consistency} is, in fact, a consequnce of QM.}
\begin{prop*} 
Let a quantum mechanical 
system be in a state, s.t. measurement of
$\hat{C}$ yields repeatable results (same result
each time). Then according to QM, \emph{weak functional consistency} holds, where $\hat{C}:=
f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n})$, and
$\hat{B}_{i}$ are as defined (in \defnref{seqnMltpl})
\end{prop*}
\begin{proof} 
Without loss of generality we can take
$\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}$ to be
mutually compatible and a complete set of
observables (operators can be added to make the set
complete if needed).
It follows
that $\exists$
$\left|\mathbf{b}=\left(b_{1},b_{2},\dots
b_{n}\right)\right\rangle $ s.t.
$\hat{B}_{i}\left|\mathbf{b}\right\rangle
=b_{i}\left|\mathbf{b}\right\rangle $,
and that
$\sum_{\mathbf{b}}\left|\mathbf{b}\right\rangle
\left\langle \mathbf{b}\right|=\hat{\mathbb{I}}$.
Let the state of the system
$\left|\psi\right\rangle $ be s.t.
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $. For
the statement to follow, one need only show that
$\left|\psi\right\rangle $ must be made of only
those $\left|\mathbf{b}\right\rangle $s, which
satisfy $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$.  This is the crucial step and
proving this is straightforward. We start with
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $ and take its inner
product with $\left\langle \mathbf{b}\right|$ to
get 
\begin{eqnarray*} 
c \langle \mathbf{b}|\psi\rangle 
& = &
\langle
\mathbf{b} |\hat{C} |\psi
\rangle  
\\
&=&
\langle
\mathbf{b}| f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}) |
\psi \rangle
\\ 
&=&
\tilde f (b_{1},b_{2},\dots b_{n})
\langle
\mathbf{b}|\psi\rangle  
\end{eqnarray*}
Also, we have $\left|\psi\right\rangle
=\sum_{\mathbf{b}}\left\langle
\mathbf{b}|\psi\right\rangle
\left|\mathbf{b}\right\rangle $, from
completeness. If we consider
$\left|\mathbf{b}\right\rangle $s for which
$\left\langle \mathbf{b}|\psi\right\rangle \neq0$,
then we find that indeed $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$. The case when $\left\langle
\mathbf{b}|\psi\right\rangle =0$ is anyway
irrelevant as the corresponding $\vert b\rangle$
does not contribute to $\vert \psi\rangle$.
We can thus conclude that
$\left|\psi\right\rangle $ is made only of those
$\left|\mathbf{b}\right\rangle $s that satisfy the
required relation.
\end{proof} 
It is worth noting that in the Peres Mermin
case, where $\hat{R}_{i}$ and $\hat{C}_{j}$ are
$\pm\hat{\mathbb{I}}$, it follows that all
states are their eigenstates. Consequently, for
these operators \emph{weak functional consistency}
must always hold.

\section{Construction}
We are now ready to describe our 
model explicitly.
Let the state of a finite dimensional quantum
system be $\left|\chi\right\rangle $. We wish to
assign a value to an arbitrary observable
$\hat{A}=\sum_{a}a\left|a\right\rangle
\left\langle a\right|$, which has eigenvectors $\{
\vert a_j\rangle \}$. The corresponding ordered
eigenvalues are $\{a_j\}$ such that $a_{\rm
min}=a_1$ and $a_{\rm max}=a_n$.  Our HV model for
QM assigns values in the following three steps:
\setdefaultleftmargin{0pt}{}{}{}{}{}
\begin{enumerate}
\item
{\bf  Initial HV:} Pick a number
$c\in[0,1]$, from a uniform random distribution.\
\item
{\bf
Assignment or Prediction:}
 The value assigned to
$\hat{A}$ is given by finding the smallest $a$
s.t.  $c\le\sum_{a'=a_{\text{min}}}^{a}\left|\left\langle
a'|\chi\right\rangle \right|^{2},$ viz. we have
specified a prediction map, $M(\hat{A},[\chi ])=a$.
\item
{\bf  Update:} After measuring an operator, the state is
updated (collapsed) in accordance with the rules
of QM. This completely specifies the sequence map
$S$.
\end{enumerate}
The above HV model works for all quantum systems,
however, to illustrate its working,
consider the example of a spin-half particle in the state
$\left|\chi\right\rangle
=\cos\theta\left|0\right\rangle
+\sin\theta\left|1\right\rangle $ and the observable
$\hat{A}=\hat{\sigma}_{z}=\left|0\right\rangle
\left\langle 0\right|-\left|1\right\rangle
\left\langle 1\right|$. Now, according to the
postulates of this theory, $M(\hat{A},[\chi])=+1$
if
the randomly generated value for 
$c\le\cos^{2}\theta$ else $\hat{A}$ is assigned
$-1$; it then follows,  from $c$ being uniformly
random in $[0,1]$ that the statistics agree with
the predictions of QM {\it i.e.}  the
Born rule. 
The assignment described by the
prediction map $M$ is non-contextual since, given
an operator and a state (alongwith the HV $c$), the value is
uniquely assigned. The map $M$ is, however,
functionally inconsistent.
